The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-03-07 Paper.Rmd
commenting out the paragraph for discussion as of now

We approached our overall goal of exploring the potential of 
copula approaches for understanding dependences in ecology by addressing the 
folling specific questions. (Q1) Do datasets in ecology and related fields 
have copula structure distinct from that of a Gaussian/normal distribution (here called 
non-normal copula structure)? Do they show tail dependence distinct from that 
of a normal distribution, and in particular do they show asymmetric tail 
dependence? (Q2) If so, what are some possible causes/mechanisms of this 
non-normal copula structure? And (Q3) what are the consequences of non-normal 
copula structure and tail dependence for our understanding of biology and for 
applications? To address Q1, we analyze several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data. We selected these datasets to span multiple 
sub-fields and organizational levels of ecology, in an attempt to be 
reasonably representative of the field as a whole. The second section of the 
paper introduces the data, and the third and fourth sections use the data to 
address Q1. The fifth and sixth sections, which address Q2, use simple models to 
explore mechanisms. The seventh and eigth sections, which address Q3, use both
the data of section two as well as models to explore a variety of consequences
of copula structure and tail dependence for applications and for our understanding
of ecology. The ninth section is the Discussion. 
Copula approaches have been used to great benefit in neuroscience [@onken2009], 
bioinformatics 
[@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], finance, and other fields. 
To our knowledge, these approaches have been used seldom or never for ecological 
research. We argue that the benefits to ecological research of a complete statistical 
desription of dependence will be substantial, and we hope this paper helps stimulate the
fruitful use of copula approaches across the discipline.

@@

The environmental, species-trait, and community datasets, as we studied them, are henceforth 
called the "bivariate" datasets in the sense that they comprise two quantities measured at 
different locations or for different species. The phenology and population datasets are 
"multivariate" in that they comprise, for each taxon, measurements at multiple locations; 
copula structure can be studied for each pair of locations. Likewise the the ecosystem 
functioning (methane flux) dataset was  multivariate in this sense.

@@

The first part Q1 was answered for a bivariate dataset 
by comparing the AIC weights of the non-normal copula
families with that of the normal family. 
The second part of Q1 was answered by comparing 
the model-averaged lower- and upper-tail dependence
values to the lower- and upper-tail dependence values of a normal copula (which 
are 0). The third part of Q1
was answered by comparing the difference of the model-averaged lower- and upper-tail 
dependence values to the analogous difference for a normal copula, again 0.
For multivariate datasets,
each of which consists of time series from multiple locations, we carried out the bivariate analysis
for all possible pairwise combinations of locations and compared AIC weights and tail-dependence of 
minimum-AIC copulas with that of the normal copula. We also counted the number of location pairs 
for which model-averaged lower-tail dependence was greater than model-averaged upper-tail dependence 
or vice-versa, and reported the percentage of pairs showing stronger lower-tail dependence out of 
the total number non-independent location pairs considered.

@@

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
bivariate data sets, the two variables were always significantly related 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC (highest-AIC-weight) copula 
(Table \ref{tab_bivar_summary}, row 2). 
The N copula was always a poor fit compared to the highest-AIC-weight copula 
(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
Upper- and lower-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
differed substantially from 0, the tail dependence statistics of 
the N copula, answering the second part of Q1; and also differed substantially from each other
(Table \ref{tab_bivar_summary}, row 9),
answering the third part of Q1. The first three data sets had stronger upper- 
than lower-tail dependence, and the last data set had the reverse. 
Although these results convincingly show non-normal copula structure, the 
model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula was a poor fit in some cases 
(Table \ref{tab_bivar_summary}, rows 5-6).
But our nonparametric statistics confirmed the same conclusions 
(Table \ref{tab_bivar_summary}, row 10).
Detailed results for each data set are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.

@@

dt <- tibble(Description = c("1. P-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC weight of minimum-AIC copula",
           "4. AIC weight of N copula",
           "5. P-value, Cramer-von Mises goodness of fit test",
           "6. P-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
knitr::kable(tab_bivar_summary, 
             format="latex",
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values in rows 5-6 are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. For row 10, the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented. If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.4cm")%>%
             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
```
<!--***DAN: Shya, it says Pl-Pu above, but I suspect it was P_first slice - P_last slice, right?
Can you please look back at the code, if necessary, and write a comment below this one 
indicating what results are actually presented in the table?
***Shya: Yes, you are right, here Pl-Pu means P_first slice - P_last slice. In table, item 10 : Assymetry rank is shown for Kendall preserving surroagates.
Reference : bivfunctionplot.R file, lines 134,135,161
***DAN: OK, great. I have made changes to deal with this, but am leaving the comment for now
as a reminder that we looked at this-->

@@

Let $E_i(t)$ be an ecological variable 
measured in location $i$ ($i=1,2$).
We initially assumed the variable follows AR(1) dynamics within a location, 
$E_i(t+1)=\beta E_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$. 
We used $\beta=0.5$. Here the $\epsilon_i(t)$ represent environmental noise; they are 
standard-normal random variables that are independent for distinct times, $t$, but 
that were made to exhibit different kinds of 
dependence across locations in different simulations (see below).
The ecological variable, $E$, is very general, and could represent, for instance, deviations
of a population density from a carrying capacity or average density, deviations of
total plant community biomass or other community variable from an average value,
or flux of a biogeochemical variable such as methane or CO$_2$..

@@

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of population densities, 
the $i^{th}$ component $P_i(t)$ representing
the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed the same 
spatial correlations (to within sampling error) for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with dispersal occurring either to nearest neighbors only,
or globally, in different simulations. The dispersal 
rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
if the population density in a patch was below $1$ (an extinction threshold), it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

<!--what we did and why is it important to study?-->
<!--***DAN: some notes by which to revise this para, Shya:
1) no need to restate methods at all, the methods are not among our main results in this paper
2) No real need to say much about why it is important to study what we stdied, that was done in the intro. Except insofar as saying what our main results are should make it clear it is an important thng to study, and see also further notes below.
3) Some additional points below
-->
Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. 
<!--***DAN: Too discursive. We do not need to repeat that we introduced this. Just summarize our main results in pothy language! And then maybe emphasize the importance, or make any additional points about importance that could not be made in the Intro. This para needs to be hard-hitting and brief, saying all the important thigs we accomplished in short form, all in one place. See below for some examples.-->
We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.
<!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well. Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results. 
<!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
among the main products of the paper - not our case. So delete the above text. If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
Our approach could help to interpret the synchronous behavior of a species from several meta-populations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that meta-populations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values. 
<!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->

<!--***DAN: Shya, the below is more summary of results. Pls summarize very briefly in the first paragraph of the Discusion ONLY. So most of the below won't be needed, but some could be incorporated into the first Discussion para above.-->
We explained the possible causes which induced asymmetric tail-dependence in the copula. Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by bivariate noise with an asymmetric copula structure showed similar tail-dependence asymmetry in the population copula from those patches (see Fig. \ref{fig_Cause4copula_params_GOF_500} and Fig. \ref{fig_Cause4copula_scatter_C_kend}). We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as the initial global noise copula (Fig. \ref{fig_ms_asym_b_0.1}). We observed significant changes in skewness of spatially averaged time series and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula. For example, Fig. \ref{fig_skewness_spearman}(A) showed green spruce aphid abundance time series had right skewed distribution but its skewness was significantly lower than was expected if it came from Spearman correlation preserving Normal surrogates. This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure. Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emissions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.

>>>>

We approached our overall goal of exploring the potential of 
copula approaches for understanding dependences in ecology by addressing the 
following specific questions. (Q1) Do datasets in ecology and related fields 
have copula structure distinct from that of a Gaussian/normal distribution (here called 
non-normal copula structure)? Do they show tail dependence distinct from that 
of a normal distribution, and in particular do they show asymmetric tail 
dependence? (Q2) If so, what are some possible causes/mechanisms of this 
non-normal copula structure? And (Q3) what are the consequences of non-normal 
copula structure and tail dependence for our understanding of biology and for 
applications? To address Q1, we analyze several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data. We selected these datasets to span multiple 
sub-fields and organizational levels of ecology, in an attempt to be 
reasonably representative of the field as a whole. The second section of the 
paper introduces the data, and the third and fourth sections use the data to 
address Q1. The fifth and sixth sections, which address Q2, use simple models to 
explore mechanisms. The seventh and eigth sections, which address Q3, use both
the data of section two as well as models to explore a variety of consequences
of copula structure and tail dependence for applications and for our understanding
of ecology. The ninth section is the Discussion. 
Copula approaches have been used to great benefit in neuroscience [@onken2009], 
bioinformatics 
[@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], finance, and other fields. 
To our knowledge, these approaches have been used seldom or never for ecological 
research. We argue that the benefits to ecological research of a complete statistical 
desription of dependence will be substantial, and we hope this paper helps stimulate the
fruitful use of copula approaches across the discipline.

@@

The environmental, species-trait, and community datasets, as we studied them, are henceforth 
called the "bivariate" datasets in the sense that they comprise two quantities measured at 
different locations or for different species. The phenology and population datasets are 
"multivariate" in that they comprise, for each taxon, measurements at multiple locations; 
copula structure can be studied for each pair of locations. Likewise the ecosystem 
functioning (methane flux) dataset was  multivariate in this sense.

@@

The first part of Q1 was answered for a bivariate dataset 
by comparing the AIC weights of the non-normal copula
families with that of the normal family. 
The second part of Q1 was answered by comparing 
the model-averaged lower- and upper-tail dependence
values to the lower- and upper-tail dependence values of a normal copula (which 
are 0). The third part of Q1
was answered by comparing the difference of the model-averaged lower- and upper-tail 
dependence values to the analogous difference for a normal copula, again 0.
For multivariate datasets,
each of which consists of time series from multiple locations, we carried out the bivariate analysis
for all possible pairwise combinations of locations and compared AIC weights and tail-dependence of 
minimum-AIC copulas with that of the normal copula. We also counted the number of location pairs 
for which model-averaged lower-tail dependence was greater than model-averaged upper-tail dependence 
or vice-versa, and reported the percentage of pairs showing stronger lower-tail dependence out of 
the total number of non-independent location pairs considered.

@@

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
bivariate data sets, the two variables were always significantly related 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC (highest-AIC-weight) copula 
(Table \ref{tab_bivar_summary}, row 2). 
The N copula was always a poor fit compared to the highest-AIC-weight copula 
(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
Lower- and upper-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
differed substantially from 0, the tail dependence statistics of 
the N copula, answering the second part of Q1; and also differed substantially from each other
(Table \ref{tab_bivar_summary}, row 9),
answering the third part of Q1. The first three data sets had stronger upper- 
than lower-tail dependence, and the last data set had the reverse. 
Although these results convincingly show non-normal copula structure, the 
model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula was a poor fit in some cases 
(Table \ref{tab_bivar_summary}, rows 5-6).
But our nonparametric statistics confirmed the same conclusions 
(Table \ref{tab_bivar_summary}, row 10).
Detailed results for each data set are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.

@@

dt <- tibble(Description = c("1. p-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC weight of minimum-AIC copula",
           "4. AIC weight of N copula",
           "5. p-value, Cramer-von Mises goodness of fit test",
           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
knitr::kable(tab_bivar_summary, 
             format="latex",
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values in rows 5-6 are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. For row 10, the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented. If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.4cm")%>%
             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
```
<!--***DAN: Shya, it says Pl-Pu above, but I suspect it was P_first slice - P_last slice, right?
Can you please look back at the code, if necessary, and write a comment below this one 
indicating what results are actually presented in the table?
***Shya: Yes, you are right, here Pl-Pu means P_first slice - P_last slice. In table, item 10 : Assymetry rank is shown for Kendall preserving surroagates.
Reference : bivfunctionplot.R file, lines 134,135,161
***DAN: OK, great. I have made changes to deal with this, but am leaving the comment for now
as a reminder that we looked at this-->

@@

Let $E_i(t)$ be an ecological variable 
measured in location $i$ ($i=1,2$).
We initially assumed the variable follows AR(1) dynamics within a location, 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$. 
We used $\beta=0.5$. Here the $\epsilon_i(t)$ represent environmental noise; they are 
standard-normal random variables that are independent for distinct times, $t$, but 
that were made to exhibit different kinds of 
dependence across locations in different simulations (see below).
The ecological variable, $E$, is very general, and could represent, for instance, deviations
of a population density from a carrying capacity or average density, deviations of
total plant community biomass or other community variable from an average value,
or flux of a biogeochemical variable such as methane or CO$_2$..

@@

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of population densities, 
the $i^{th}$ component $P_i(t)$ representing
the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed the same 
spatial correlations (to within sampling error) for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with dispersal occurring either to nearest neighbors only,
or globally, in different simulations. The dispersal 
rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
if the population density in a patch was below $1$ (an extinction threshold), it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations $p_0 = 50$ in each patch of $N$, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

<!--what we did and why is it important to study?-->
<!--***DAN: some notes by which to revise this para, Shya:
1) no need to restate methods at all, the methods are not among our main results in this paper
2) No real need to say much about why it is important to study what we stdied, that was done in the intro. Except insofar as saying what our main results are should make it clear it is an important thng to study, and see also further notes below.
3) Some additional points below
-->
<!--
Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. -->
<!--***DAN: Too discursive. We do not need to repeat that we introduced this. Just summarize our main results in pothy language! And then maybe emphasize the importance, or make any additional points about importance that could not be made in the Intro. This para needs to be hard-hitting and brief, saying all the important thigs we accomplished in short form, all in one place. See below for some examples.-->

<!--We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.-->

<!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well. Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
<!--We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results. -->
<!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
among the main products of the paper - not our case. So delete the above text. If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
<!--Our approach could help to interpret the synchronous behavior of a species from several meta-populations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that meta-populations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values. -->
<!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->

<!--***DAN: Shya, the below is more summary of results. Pls summarize very briefly in the first paragraph of the Discusion ONLY. So most of the below won't be needed, but some could be incorporated into the first Discussion para above.-->
<!--We explained the possible causes which induced asymmetric tail-dependence in the copula. Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by bivariate noise with an asymmetric copula structure showed similar tail-dependence asymmetry in the population copula from those patches (see Fig. \ref{fig_Cause4copula_params_GOF_500} and Fig. \ref{fig_Cause4copula_scatter_C_kend}). We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as the initial global noise copula (Fig. \ref{fig_ms_asym_b_0.1}). We observed significant changes in skewness of spatially averaged time series and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula. For example, Fig. \ref{fig_skewness_spearman}(A) showed green spruce aphid abundance time series had right skewed distribution but its skewness was significantly lower than was expected if it came from Spearman correlation preserving Normal surrogates. This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure. Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emissions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment. 
-->
